\documentclass{article}\usepackage{amsmath}\begin{document}\title{Simplex Solver}\maketitle\begin{flushleft}\textbf{Problem}\end{flushleft}\begin{flushleft}Given the following linear system and objective function, find the optimal solution.\end{flushleft}\begin{equation*}\max{ x_1 + 3x_2 + 2x_3 } \\ \end{equation*}\[\left\{\begin{array}{c} 2x_1 + 4x_2 + x_3 \leq -6 \\ - x_1 + 6x_2 + x_3 \leq 1 \\  x_1 + 8x_2 + 3x_3 \leq 2 \\ \end{array}\right.\]\begin{flushleft}\textbf{Solution}\end{flushleft}\begin{flushleft}Add slack variables to turn all inequalities to equalities.\end{flushleft}\[\left\{\begin{array}{c} 2x_1 + 4x_2 + x_3 + s_1 = -6 \\ - x_1 + 6x_2 + x_3 + s_2 = 1 \\  x_1 + 8x_2 + 3x_3 + s_3 = 2 \\ \end{array}\right.\]\begin{flushleft}Create the initial tableau of the new linear system.\end{flushleft}\begin{equation*}\begin{bmatrix}\begin{array}{cccccc|c}x_1 &x_2 &x_3 &s_1 &s_2 &s_3 &b \\ \hline2 & 4 & 1 & 1 & 0 & 0 & -6 \\-1 & 6 & 1 & 0 & 1 & 0 & 1 \\1 & 8 & 3 & 0 & 0 & 1 & 2 \\ \hline-1 & -3 & -2 & 0 & 0 & 0 & 0 \\\end{array}\end{bmatrix}\begin{array}{c}\\s_1 \\s_2 \\s_3 \\\\\end{array}\end{equation*}\begin{flushleft}There are negative elements in the bottom row, so the current solution is not optimal. Thus, pivot to improve the current solution. The entering variable is $x_2$ and the departing variable is $s_2$.\end{flushleft}\begin{flushleft}Perform elementary row operations until the pivot element is 1 and all other elements in the entering column are 0.\end{flushleft}\begin{equation*}\begin{bmatrix}\begin{array}{cccccc|c}x_1 &x_2 &x_3 &s_1 &s_2 &s_3 &b \\ \hline8/3 & 0 & 1/3 & 1 & -2/3 & 0 & -20/3 \\-1/6 & 1 & 1/6 & 0 & 1/6 & 0 & 1/6 \\7/3 & 0 & 5/3 & 0 & -4/3 & 1 & 2/3 \\ \hline-3/2 & 0 & -3/2 & 0 & 1/2 & 0 & 1/2 \\\end{array}\end{bmatrix}\begin{array}{c}\\s_1 \\x_2 \\s_3 \\\\\end{array}\end{equation*}\begin{flushleft}There are negative elements in the bottom row, so the current solution is not optimal. Thus, pivot to improve the current solution. The entering variable is $x_1$ and the departing variable is $s_3$.\end{flushleft}\begin{flushleft}Perform elementary row operations until the pivot element is 1 and all other elements in the entering column are 0.\end{flushleft}\begin{equation*}\begin{bmatrix}\begin{array}{cccccc|c}x_1 &x_2 &x_3 &s_1 &s_2 &s_3 &b \\ \hline0 & 0 & -11/7 & 1 & 6/7 & -8/7 & -52/7 \\0 & 1 & 2/7 & 0 & 1/14 & 1/14 & 3/14 \\1 & 0 & 5/7 & 0 & -4/7 & 3/7 & 2/7 \\ \hline0 & 0 & -3/7 & 0 & -5/14 & 9/14 & 13/14 \\\end{array}\end{bmatrix}\begin{array}{c}\\s_1 \\x_2 \\x_1 \\\\\end{array}\end{equation*}\begin{flushleft}There are negative elements in the bottom row, so the current solution is not optimal. Thus, pivot to improve the current solution. The entering variable is $x_3$ and the departing variable is $x_1$.\end{flushleft}\begin{flushleft}Perform elementary row operations until the pivot element is 1 and all other elements in the entering column are 0.\end{flushleft}\begin{equation*}\begin{bmatrix}\begin{array}{cccccc|c}x_1 &x_2 &x_3 &s_1 &s_2 &s_3 &b \\ \hline11/5 & 0 & 0 & 1 & -2/5 & -1/5 & -34/5 \\-2/5 & 1 & 0 & 0 & 3/10 & -1/10 & 1/10 \\7/5 & 0 & 1 & 0 & -4/5 & 3/5 & 2/5 \\ \hline3/5 & 0 & 0 & 0 & -7/10 & 9/10 & 11/10 \\\end{array}\end{bmatrix}\begin{array}{c}\\s_1 \\x_2 \\x_3 \\\\\end{array}\end{equation*}\begin{flushleft}There are negative elements in the bottom row, so the current solution is not optimal. Thus, pivot to improve the current solution. The entering variable is $s_2$ and the departing variable is $x_2$.\end{flushleft}\begin{flushleft}Perform elementary row operations until the pivot element is 1 and all other elements in the entering column are 0.\end{flushleft}\begin{equation*}\begin{bmatrix}\begin{array}{cccccc|c}x_1 &x_2 &x_3 &s_1 &s_2 &s_3 &b \\ \hline5/3 & 4/3 & 0 & 1 & 0 & -1/3 & -20/3 \\-4/3 & 10/3 & 0 & 0 & 1 & -1/3 & 1/3 \\1/3 & 8/3 & 1 & 0 & 0 & 1/3 & 2/3 \\ \hline-1/3 & 7/3 & 0 & 0 & 0 & 2/3 & 4/3 \\\end{array}\end{bmatrix}\begin{array}{c}\\s_1 \\s_2 \\x_3 \\\\\end{array}\end{equation*}\begin{flushleft}There are negative elements in the bottom row, so the current solution is not optimal. Thus, pivot to improve the current solution. The entering variable is $x_1$ and the departing variable is $x_3$.\end{flushleft}\begin{flushleft}Perform elementary row operations until the pivot element is 1 and all other elements in the entering column are 0.\end{flushleft}\begin{equation*}\begin{bmatrix}\begin{array}{cccccc|c}x_1 &x_2 &x_3 &s_1 &s_2 &s_3 &b \\ \hline0 & -12 & -5 & 1 & 0 & -2 & -10 \\0 & 14 & 4 & 0 & 1 & 1 & 3 \\1 & 8 & 3 & 0 & 0 & 1 & 2 \\ \hline0 & 5 & 1 & 0 & 0 & 1 & 2 \\\end{array}\end{bmatrix}\begin{array}{c}\\s_1 \\s_2 \\x_1 \\\\\end{array}\end{equation*}\begin{flushleft}There are no negative elements in the bottom row, so we know the solution is optimal. Thus, the solution is: \end{flushleft}\begin{equation*}s_1 = -10, s_2 = 3, s_3 = 0, x_1 = 2, x_2 = 0, x_3 = 0, z = 2\end{equation*}\end{document}